# IODS course chapter 2 Regression analysis 

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using  

To perform linear regression analysis in R, the first step is to make sure that the data are in properly formatted in a dataframe. This has already been performed. Now the first step is to bring the data here so that I can start working on it. I have moved the csv file I created earlier in this working directory so I can read it directly. I will call the new dataframe object df:


```{r}
df <- read.csv("learning2014.csv")

```

Function str() shows the structure of the dataframe and tells that it has 166 observations od 7 variables and function head() shows the first six rows of the dataframe. This way I can be sure that the data is correctly imported. 
```{r}
str(df)
head(df)
```

summary() function prints a summary of each variable.
```{r}

summary(df)
```





